New Metrics for Assessing Projection Pursuit Indexes, and Guiding Optimisation Choices

### 2024-09-12 jso.qmd
small edits on the Practicalities section

Smoothness and squintability are computed following the procedures
outlined in @sec-smoothness and @sec-squintability and as illustrated in
@fig-smoothness and @fig-squintability.

To compute smoothness, 300 random bases are simulated. Index values and
projection distance (to the optimal basis) are calculated for each
random basis before fitting the Gaussian process model to obtain the
smoothness measure for the index.

@@

The bases visited by jellyfish No. 1 is extracted with `filter(loop = 1)` (the variable `loop` in the data object is the jellyfish ID) and the `check_dup()` function removes the bases in the iteration without improvements. The bases are then supplied as a `planned_tour()` in `animate_xy()` to visualise the tour path of jellyfish No.1.

@@

To sample bases for calculating smoothness, the required arguments include `idx` (the index function used), `data` (the dataset), `n_basis` (the number of random bases to sample), and `best` (the best basis to compute projection distance). The output of `sample_bases()` is a data frame with four columns: `id` (the basis ID), `basis` (the basis matrix sampled), `dist` (the projection distance to the best basis), and `index` (the index value). To speed up the computation of index values,  parallellasation is available through the `parallel = TRUE` argument.

The `calc_smoothness()` function takes the output from `sample_bases()` and fits a Gaussian process model to the index values against the projection distances to compute the smoothness metric. The starting parameter for the fit can be specified using the `start_params` argument, and additional Gaussian process arguments can be provided through `other_gp_params`. The smoothness metric is the smoothness parameter from the Gussian process and all details of the fit can be accessed using `attr(sm_res, "fit_res")`.

@@

The `calc_squintability()` function computes squintability using two methods: 1) parametrically, by fitting a scaled sigmoid function through non-linear least square (`method = "nls"`), and 2) non-parametrically, using kernel smoothing (`method = "ks"`). A `bin_width` argument is required to average the index values over the projection distance before the fit. The output provides the estimated parameters for the scaled sigmoid function ($\theta_1$ to $\theta_4$) and the squintability metric for the parametric case. For the non-parametric case, it shows the (`max_d`), the corresponding projection distance (`max_dist`), and teh squintability metric as their products.

>>>>

Smoothness and squintability are computed following the procedures
outlined in @sec-smoothness and @sec-squintability and as illustrated in
@fig-smoothness and @fig-squintability. To compute smoothness, 300 random
bases are simulated. Index values and
projection distance (to the optimal basis) are calculated for each
random basis before fitting the Gaussian process model to obtain the
smoothness measure for the index.

@@

@@

To sample bases for calculating smoothness, four arguments are required: 1) the index function, the dataset, the number of random bases to sample, and the best basis to compute projection distance. The output of `sample_bases()` is a data frame containing columns/ list-columns of sampled basis matrix, calculated projection distance and index value. Parallellasation is available through the `parallel = TRUE` argument to speed up the computation.

The `calc_smoothness()` function takes the output from `sample_bases()` and fits a Gaussian process model to the index values against the projection distances to compute the smoothness metric. Starting parameters and additional Gaussian process arguments can be specified and details of the fit can be accessed through the `fit_res` attribute of the `calc_smoothness()` output.

@@

The `calc_squintability()` function computes squintability using two methods: 1) parametrically, by fitting a scaled sigmoid function through non-linear least square (`method = "nls"`), and 2) non-parametrically, using kernel smoothing (`method = "ks"`). A `bin_width` argument is required to average the index values over the projection distance before the fit. The output provides the estimated parameters for the scaled sigmoid function ($\theta_1$ to $\theta_4$) and the squintability metric for the parametric case. For the non-parametric case, it shows the maximum gradient attained(`max_d`), the corresponding projection distance (`max_dist`), and the squintability metric as their products.

### 2024-09-12 jso.tex
small edits on the Practicalities section

\begin{tcolorbox}[enhanced jigsaw, opacitybacktitle=0.6, bottomtitle=1mm, coltitle=black, title={Algorithm: Jellyfish Optimizer Pseudo Code}, colback=white, breakable, toptitle=1mm, colframe=quarto-callout-note-color-frame, leftrule=.75mm, left=2mm, bottomrule=.15mm, rightrule=.15mm, toprule=.15mm, colbacktitle=quarto-callout-note-color!10!white, arc=.35mm, titlerule=0mm, opacityback=0]

@@

Smoothness and squintability are computed following the procedures
outlined in Section~\ref{sec-smoothness} and
Section~\ref{sec-squintability} and as illustrated in
Figure~\ref{fig-smoothness} and Figure~\ref{fig-squintability}.

To compute smoothness, 300 random bases are simulated. Index values and
projection distance (to the optimal basis) are calculated for each
random basis before fitting the Gaussian process model to obtain the
smoothness measure for the index.

@@

\centering

@@

The bases visited by jellyfish No.~1 is extracted with
\texttt{filter(loop\ =\ 1)} (the variable \texttt{loop} in the data
object is the jellyfish ID) and the \texttt{check\_dup()} function
removes the bases in the iteration without improvements. The bases are
then supplied as a \texttt{planned\_tour()} in \texttt{animate\_xy()} to
visualise the tour path of jellyfish No.1.

@@

To sample bases for calculating smoothness, the required arguments
include \texttt{idx} (the index function used), \texttt{data} (the
dataset), \texttt{n\_basis} (the number of random bases to sample), and
\texttt{best} (the best basis to compute projection distance). The
output of \texttt{sample\_bases()} is a data frame with four columns:
\texttt{id} (the basis ID), \texttt{basis} (the basis matrix sampled),
\texttt{dist} (the projection distance to the best basis), and
\texttt{index} (the index value). To speed up the computation of index
values, parallellasation is available through the
\texttt{parallel\ =\ TRUE} argument.

The \texttt{calc\_smoothness()} function takes the output from
\texttt{sample\_bases()} and fits a Gaussian process model to the index
values against the projection distances to compute the smoothness
metric. The starting parameter for the fit can be specified using the
\texttt{start\_params} argument, and additional Gaussian process
arguments can be provided through \texttt{other\_gp\_params}. The
smoothness metric is the smoothness parameter from the Gussian process
and all details of the fit can be accessed using
\texttt{attr(sm\_res,\ "fit\_res")}.

@@

The \texttt{calc\_squintability()} function computes squintability using
two methods: 1) parametrically, by fitting a scaled sigmoid function
through non-linear least square (\texttt{method\ =\ "nls"}), and 2)
non-parametrically, using kernel smoothing (\texttt{method\ =\ "ks"}). A
\texttt{bin\_width} argument is required to average the index values
over the projection distance before the fit. The output provides the
estimated parameters for the scaled sigmoid function (\(\theta_1\) to
\(\theta_4\)) and the squintability metric for the parametric case. For
the non-parametric case, it shows the (\texttt{max\_d}), the
corresponding projection distance (\texttt{max\_dist}), and teh
squintability metric as their products.

>>>>

\begin{tcolorbox}[enhanced jigsaw, rightrule=.15mm, breakable, toprule=.15mm, bottomtitle=1mm, colframe=quarto-callout-note-color-frame, arc=.35mm, titlerule=0mm, coltitle=black, bottomrule=.15mm, opacityback=0, opacitybacktitle=0.6, title={Algorithm: Jellyfish Optimizer Pseudo Code}, colbacktitle=quarto-callout-note-color!10!white, left=2mm, colback=white, leftrule=.75mm, toptitle=1mm]

@@

Smoothness and squintability are computed following the procedures
outlined in Section~\ref{sec-smoothness} and
Section~\ref{sec-squintability} and as illustrated in
Figure~\ref{fig-smoothness} and Figure~\ref{fig-squintability}. To
compute smoothness, 300 random bases are simulated. Index values and
projection distance (to the optimal basis) are calculated for each
random basis before fitting the Gaussian process model to obtain the
smoothness measure for the index.

@@

[H] \centering

@@

@@

To sample bases for calculating smoothness, four arguments are required:
1) the index function, the dataset, the number of random bases to
sample, and the best basis to compute projection distance. The output of
\texttt{sample\_bases()} is a data frame containing columns/
list-columns of sampled basis matrix, calculated projection distance and
index value. Parallellasation is available through the
\texttt{parallel\ =\ TRUE} argument to speed up the computation.

The \texttt{calc\_smoothness()} function takes the output from
\texttt{sample\_bases()} and fits a Gaussian process model to the index
values against the projection distances to compute the smoothness
metric. Starting parameters and additional Gaussian process arguments
can be specified and details of the fit can be accessed through the
\texttt{fit\_res} attribute of the \texttt{calc\_smoothness()} output.

@@

The \texttt{calc\_squintability()} function computes squintability using
two methods: 1) parametrically, by fitting a scaled sigmoid function
through non-linear least square (\texttt{method\ =\ "nls"}), and 2)
non-parametrically, using kernel smoothing (\texttt{method\ =\ "ks"}). A
\texttt{bin\_width} argument is required to average the index values
over the projection distance before the fit. The output provides the
estimated parameters for the scaled sigmoid function (\(\theta_1\) to
\(\theta_4\)) and the squintability metric for the parametric case. For
the non-parametric case, it shows the maximum gradient
attained(\texttt{max\_d}), the corresponding projection distance
(\texttt{max\_dist}), and the squintability metric as their products.
